When are Fuzzy Extractors Possible?

### 2014-10-06 trunk/info-bounds/FuzzyExtPossibility.tex
minor edits

Let $V$ be the process of uniformly sampling $W\leftarrow \mathcal{W}$ and then sampling $w\leftarrow W$.  Let $Z$ be the description of which $W$ was sampled.  The view of the challenger is $V$, while the view of the adversary is a distribution $V|Z$ where $Z$ is the description of which $W$ was sampled.  Our results rule out security for an average member of $\mathcal{W}$.  It may be possible to improve parameters by ruling out only a worst case $W$.  In \apref{sec:def equiv}, we show that providing security for a family $\mathcal{W}$ is equivalent to providing security for all distributions over that family.

@@

\begin{theorem}
\label{thm:imposs sketch}
Let $n$ be a security parameter.  There exists a family of distributions $\mathcal{W}$ such that for each element $W\in \mathcal{W}$, $\Hfuzz(W)= \omega(\log n)$, for any $(\mathcal{M}, \mathcal{W}, \tilde{m}, t)$-secure sketch with error $\delta <1/4$ and $t\ge 4$ then $\tilde{m}<2$.

Furthermore, this is true on average.  Let $V$ be process of uniformly sampling $W\leftarrow \mathcal{W}$ and sampling $w\leftarrow W$ and let $Z$ be the description of which $W$ is sampled.  Then, 
\[
\Hav(V|\sketch(V), Z)< 2.
\]
\end{theorem}

@@

Secure sketches that provide computational unpredictability are implied by virtual-grey box obfuscation of proximity functions~\cite{BitanskyCKP14}.  Our impossibility result says nothing about this weaker form of a secure sketch.  Extracting from unpredictability requires a extractor with a reconstruction property~\cite{barak-computational, DBLP:conf/eurocrypt/HsiaoLR07}.

@@

\begin{theorem}
\label{thm:imposs fuzz ext}
Let $n$ be a security parameter.  There exists a family of distributions $\mathcal{W}$ over $\zo^n$ such that for each element $W\in \mathcal{W}$, $\Hfuzz(W)= \omega(\log n)$.  Let $\kappa \ge 2$, $t = \omega(n^{1/2}\log n)$.  Any $(\mathcal{M}, \mathcal{W}, \kappa, t, \epsilon)$-fuzzy extractor with error $\delta = 0$ has $\epsilon > 1/8 - \ngl(n)$.

Furthermore, this is true on average.  Let $V$ be process of uniformly sampling $W\leftarrow \mathcal{W}$ and sampling $w\leftarrow W$ and let $Z$ be the description of which $W$ is sampled.  Let $(\Key, P)\leftarrow \gen(V)$.  Then, 
\[
\sd ((\Key,P, Z), (U_\kappa, P, Z))>1/8-\ngl(n).
\]

@@

Our counterexample uses a slightly different family of distributions $\mathcal{W}$ than the counterexample for secure sketches.  
We can't work directly on the Hamming distance over a large alphabet~(as we did for secure sketches).  We embed this larger space to the binary Hamming metric.
Let $x_1,..., x_\nu \in \zo^\nu$.  Let $\mathbb{F}$ denote the field of size $2^{\nu}$.  Let $a_2,..., a_{n/\nu}\in\mathbb{F}$ such that $a_i\neq 0$ and let $b_2,..., b_{n/\nu}\in\mathbb{F}$.  
Interpret $x_1,..., x_{\nu}$ as a element $x\in \mathbb{F}$ and let 
\[w =  \begin{pmatrix} \vec{1} \\a_2  \\ \vdots \\ a_{n/\nu} \end{pmatrix} x + \begin{pmatrix} 0  \\ 
b_2\\ \vdots \\ b_{n/\nu} \end{pmatrix} .
\]
The multiplication is in $\mathbb{F}$.
Define a distribution $W$ as the uniform distribution over values of $x$ for a particular value of $a_2,..., a_{n/\nu}$, $b_2,..., b_{n/\nu}$.  
Let $\mathcal{W}$ be the set of all such $W$.

 Define $V$ as the process of uniformly choosing $W\leftarrow \mathcal{W}$ and then sampling from $w\leftarrow W$.  The adversary sees $\sketch(V)$ and which distribution was sampled, call this information $Z$. $Z$ is the description of the line $Z = a_2,..., a_{n/\nu}, b_2, ..., b_{n/\nu}$.

@@

The best strategy for an adversary that receives $i$ as is to guess a point that has the most nearby weight in that level.  The adversary chooses \[w^*= \argmax_{w' \in \mathcal{M}}\Pr_{w\in W | 2^{-(i+1)}< \Pr[W=w]\le 2^{-i}\wedge \dis(w, w^*)}[W=w].\] The success of this adversary is at least $2^{-(i+1)}|B_{t,i, max}|$ as there at $B_{t, i, max}$ nearby points in that layer each with probability at least $2^{-(i+1)}$.  There are $n$ outcomes for $i$. The overall success of such an adversary is at most $n$ better than an adversary without such input~(by~\cite[Lemma 2.2]{DBLP:journals/siamcomp/DodisORS08}).  That is, 
\begin{align}
\expe_{i | m\le i \le m+n}2^{-(i+1)}|B_{t, i, max}|&\le \expe_{i | m\le i \le n+m}\left( \max_{w^*\in W}\sum_{w\in W| 2^{-(i+1)}< \Pr[W=w]\le 2^{-i} \wedge \dis(w, w^*) \le t}\Pr [W=w]\right) \nonumber\\
&\le n \left(\max_{w^*\in W} \sum_{w\in W | \dis(w, w^*)\le t} \Pr[W=w]\right)\nonumber \\
&= n 2^{-\Hfuzz(W)}\label{eq:link fuzz 3}
\end{align}
\textbf{Security:}
We now argue security.  First note that the total weight of points whose probability is less than $2^{-(n+m)}$ is at most $2^{-m}$~(there are at most $2^n$ points in the distribution).  Let $1_{\text{low}}$ be the indicator random variable for $\Pr[W=w]\le 2^{-(n+m)}$.  Then 
\begin{align*}
\Hav(W | \sketch(W)) = -\log \left(\Pr[1_{\text{low}}=1] * 1 + \Pr[1_{\text{low}} =0]   2^{-\Hav(W | \sketch(W) \wedge 1_{\text{low}} = 0)}\right)\\
-\log\left( 2^{-m} + (1-2^{-m})2^{-\Hav(W | \sketch(W) \wedge 1_{\text{low}} = 0)}\right)
\end{align*}
For the remainder of the proof, we seek a bound on 
\[
2^{-\Hav(W | \sketch(W) \wedge 1_{\text{low}} =0} = \max_{w\in W | 2^{-(n+m)}<\Pr[W=w]}\Pr[W=w | \sketch(W)].
\]
We separate out this quantity into levels:
\begin{align*}
\max_{w\in W | \Pr[W=w]>2^{-(m+n)}}\left(\Pr[W=w | \sketch(W)]\right) &= \expe_{i | m\le i \le m+n} \left(\max_{w\in W | \Pr[W=w]\in L_i} \Pr[W=w | \sketch(W), i]\right)\\
&= \expe_{i | m\le i \le m+n} \left(\max_{w\in W | \Pr[W=w]\in L_i} \Pr[W=w]*2^{|\sketch(W)|i|}\right)\\
&\le \expe_{i | m\le i \le m+n} \left(\max_{w\in W | \Pr[W=w]\in L_i} \Pr[W=w]*2^{H_0(\sketch(W) | i)}\right)\\
&\le \expe_{i | m\le i \le m+n} \left(2^{-i}*|B_{t, i, max}|/\delta\right)\\
&\le\frac{ \expe_{i | m\le i \le m+n} \left(2^{-(i+1)}*|B_{t, i, max}|\right)}{2\delta}\\
%&=\delta \sum_{i | m\le i \le m+n} \Pr[W\in L_i]\left(2^{-i}*|B_{t, i, max}|\right)
%\end{align*}
%Now consider This allows us to conclude that 
%\begin{align*}
%\max_{w\in W | \Pr[W=w]>2^{-n}\epsilon}\left(\Pr[W=w | \sketch(W)]\right) 
&= \frac{n 2^{-\Hfuzz(W)}}{2\delta}.
\end{align*}
Where the last line follows by Equation~\eqref{eq:link fuzz 3}.
Combining both cases we have:
\begin{align*}
\Hav(W | \sketch(W)) &= -\log \left(2^{-m}+\frac{(1-2^{-m})(n)2^{-\Hfuzz(W)}}{2\delta}\right)\\
&\ge -\log \min\{2^{-m}, \frac{(1-2^{-m}) n2^{-\Hfuzz(W)}}{2\delta}\})-1\\
&\ge \Hfuzz(W) - \log n + \log \delta - \log (1-2^{-m}) - 2\\
&\ge \Hfuzz(W) - \log n + \log \delta - 3\\
\end{align*}
Where the third line follows from the second because $\Hfuzz(W)\le \Hoo(W) = m$. The last line follows from the fourth because if $m\ge 1$ then $\log (1-2^{-m})\le 1$ and if $m< 1$ the entire bound is vacuous as $\Hfuzz(W)< 1$.
\end{proof}

>>>>

Let $V$ be the process of uniformly sampling $W\leftarrow \mathcal{W}$ and then sampling $w\leftarrow W$.  Let the random variable $Z$ indicate which $W$ was sampled.  The view of the challenger is $V$, while the view of the adversary is a distribution $V|Z$.  Our results rule out security for an average member of $\mathcal{W}$.  It may be possible to improve parameters by ruling out only a worst case $W$.  In \apref{sec:def equiv}, we show that providing security for a family $\mathcal{W}$ is equivalent to providing security for all distributions over that family.

@@

\begin{theorem}
\label{thm:imposs sketch}
Let $n$ be a security parameter.  There exists a family of distributions $\mathcal{W}$ such that for each element $W\in \mathcal{W}$, $\Hfuzz(W)= \omega(\log n)$, and yet for any $(\mathcal{M}, \mathcal{W}, \tilde{m}, t)$-secure sketch $(\sketch, \rec)$ with error $\delta <1/4$ and distance $t\ge 4$, the remaining entropy $\tilde{m}<2$.

Furthermore, this is true on average.  Let $V$ be process of uniformly sampling $W\leftarrow \mathcal{W}$ and sampling $w\leftarrow W$, and let $Z$ indicate which $W$ is sampled.  Then
\[
\Hav(V|\sketch(V), Z)< 2.
\]
\end{theorem}

@@

Secure sketches that provide computational unpredictability are implied by virtual-grey box obfuscation of proximity functions~\cite{BitanskyCKP14}.  Our impossibility result says nothing about this weaker form of a secure sketch.  Extraction from unpredictability entropy can be done using an extractor with a reconstruction property~\cite{barak-computational, DBLP:conf/eurocrypt/HsiaoLR07}; however, in the computational setting, the obfuscated function can simply hide a randomly generated key, and therefore extraction is not necessary to obtain a fuzzy extractor.

@@

\begin{theorem}
\label{thm:imposs fuzz ext}
Let $n$ be a security parameter.  There exists a family of distributions $\mathcal{W}$ over $\zo^n$ satisfying the following conditions. For each element $W\in \mathcal{W}$, $\Hfuzz(W)= \omega(\log n)$. Let $\kappa \ge 2$ and $t = \omega(n^{1/2}\log n)$.  Any $(\mathcal{M}, \mathcal{W}, \kappa, t, \epsilon)$-fuzzy extractor with error $\delta = 0$ has $\epsilon > 1/8 - \ngl(n)$.

Furthermore, this is true on average.  Let $V$ be process of uniformly sampling $W\leftarrow \mathcal{W}$ and sampling $w\leftarrow W$ and let $Z$ indicate which $W$ is sampled.  Let $(\Key, P)\leftarrow \gen(V)$.  Then, 
\[
\sd ((\Key,P, Z), (U_\kappa, P, Z))>1/8-\ngl(n)\,.
\]

@@

Our counterexample uses a slightly different family of distributions $\mathcal{W}$ than the counterexample for secure sketches.  
We can't work directly on the Hamming distance over a large alphabet~(as we did for secure sketches). \lnote{it would be good to explain why not}  We embed this larger space to the binary Hamming metric.
Let $x_1,..., x_\nu \in \zo^\nu$.  Let $\mathbb{F}$ denote the field of size $2^{\nu}$.  Let $a_2,..., a_{n/\nu}\in\mathbb{F}$ such that $a_i\neq 0$ and let $b_2,..., b_{n/\nu}\in\mathbb{F}$.  
Interpret $x_1,..., x_{\nu}$ as a element $x\in \mathbb{F}$ and let 
\[w =  \begin{pmatrix} \vec{1} \\a_2  \\ \vdots \\ a_{n/\nu} \end{pmatrix} x + \begin{pmatrix} 0  \\ 
b_2\\ \vdots \\ b_{n/\nu} \end{pmatrix} .
\]
The multiplication is in $\mathbb{F}$.
Define a distribution $W$ as the uniform distribution over values of $x$ for a particular value of $a_2,..., a_{n/\nu}$, $b_2,..., b_{n/\nu}$.  
Let $\mathcal{W}$ be the set of all such $W$.

 Define $V$ as the process of uniformly choosing $W\leftarrow \mathcal{W}$ and then sampling from $w\leftarrow W$.  The adversary sees $\sketch(V)$  and $Z$, where $Z$ is the description of the line $Z = a_2,..., a_{n/\nu}, b_2, ..., b_{n/\nu}$.

@@

The best strategy for an adversary that receives $i$ as is to guess a point that has the most nearby weight in that level.  The adversary chooses \[w^*= \argmax_{w' \in \mathcal{M}}\Pr_{w\in W | 2^{-(i+1)}< \Pr[W=w]\le 2^{-i}\wedge \dis(w, w^*)}[W=w].\] The success of this adversary is at least $2^{-(i+1)}|B_{t,i, max}\cap\supp(W)|$ as there at $|B_{t, i, max}\cap\supp(W)|$ nearby points in that layer each with probability at least $2^{-(i+1)}$.  There are $n$ outcomes for $i$. The overall success of such an adversary is at most $n$ better than an adversary without such input~(by~\cite[Lemma 2.2]{DBLP:journals/siamcomp/DodisORS08}).  That is, 
\begin{align}
\expe_{i | m\le i \le m+n}2^{-(i+1)}|B_{t, i, max}\cap\supp(W)|&\le \expe_{i | m\le i \le n+m}\left( \max_{w^*\in W}\sum_{w\in W| 2^{-(i+1)}< \Pr[W=w]\le 2^{-i} \wedge \dis(w, w^*) \le t}\Pr [W=w]\right) \nonumber\\
&\le n \left(\max_{w^*\in W} \sum_{w\in W | \dis(w, w^*)\le t} \Pr[W=w]\right)\nonumber \\
&= n 2^{-\Hfuzz(W)}\label{eq:link fuzz 3}
\end{align}
\textbf{Security:}
We now argue security.  First note that the total weight of points whose probability is less than $2^{-(n+m)}$ is at most $2^{-m}$~(there are at most $2^n$ points in the distribution).  Let $1_{\text{low}}$ be the indicator random variable for $\Pr[W=w]\le 2^{-(n+m)}$.  Then 
\begin{align*}
\Hav(W | \sketch(W)) = -\log \left(\Pr[1_{\text{low}}=1] * 1 + \Pr[1_{\text{low}} =0]   2^{-\Hav(W | \sketch(W) \wedge 1_{\text{low}} = 0)}\right)\\
-\log\left( 2^{-m} + (1-2^{-m})2^{-\Hav(W | \sketch(W) \wedge 1_{\text{low}} = 0)}\right)
\end{align*}
For the remainder of the proof, we seek a bound on 
\[
2^{-\Hav(W | \sketch(W) \wedge 1_{\text{low}} =0} = \max_{w\in W | 2^{-(n+m)}<\Pr[W=w]}\Pr[W=w | \sketch(W)].
\]
We separate out this quantity into levels:
\begin{align*}
\max_{w\in W | \Pr[W=w]>2^{-(m+n)}}\left(\Pr[W=w | \sketch(W)]\right) &= \expe_{i | m\le i \le m+n} \left(\max_{w\in W | \Pr[W=w]\in L_i} \Pr[W=w | \sketch(W), i]\right)\\
&= \expe_{i | m\le i \le m+n} \left(\max_{w\in W | \Pr[W=w]\in L_i} \Pr[W=w]\cdot 2^{|\sketch(W)|i|}\right)\\
&\le \expe_{i | m\le i \le m+n} \left(\max_{w\in W | \Pr[W=w]\in L_i} \Pr[W=w]\cdot 2^{H_0(\sketch(W) | i)}\right)\\
&\le \expe_{i | m\le i \le m+n} \left(2^{-i}*|B_{t, i, max}\cap\supp(W)|/\delta\right)\\
&\le\frac{ \expe_{i | m\le i \le m+n} \left(2^{-(i+1)}\cdot |B_{t, i, max}\cap \supp(W)|\right)}{2\delta}\\
%&=\delta \sum_{i | m\le i \le m+n} \Pr[W\in L_i]\left(2^{-i}*|B_{t, i, max}|\right)
%\end{align*}
%Now consider This allows us to conclude that 
%\begin{align*}
%\max_{w\in W | \Pr[W=w]>2^{-n}\epsilon}\left(\Pr[W=w | \sketch(W)]\right) 
&= \frac{n 2^{-\Hfuzz(W)}}{2\delta}.
\end{align*}
Where the last line follows by Equation~\eqref{eq:link fuzz 3}.
Combining both cases we have:
\begin{align*}
\Hav(W | \sketch(W)) &= -\log \left(2^{-m}+\frac{(1-2^{-m})(n)2^{-\Hfuzz(W)}}{2\delta}\right)\\
&\ge -\log \min\{2^{-m}, \frac{(1-2^{-m}) n2^{-\Hfuzz(W)}}{2\delta}\})-1\\
&\ge \Hfuzz(W) - \log n + \log \delta - \log (1-2^{-m}) - 2\\
&\ge \Hfuzz(W) - \log n + \log \delta - 3\\
\end{align*}
Where the third line follows from the second because $\Hfuzz(W)\le \Hoo(W) = m$. The last line follows from the fourth because if $m\ge 1$ then $\log (1-2^{-m})\le 1$ and if $m< 1$ the entire bound is vacuous as $\Hfuzz(W)< 1$.
\end{proof}
